Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector

### 2018-08-11 documentation/paper.tex
Further clarified semi-supervised stuff, and scrapped misleading graphs

However, the lack of decisiveness observed even when accuracy was relatively high, as seen in Figure \ref{waveform_best_validation}, implies that without $WallCut(D)$, this is a suboptimal technique.

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{waveform_best_validation}
    \caption{\label{waveform_best_validation} Accuracy of $DeepConvWithPos(\omega, \chi)$ compared to Acoustic Parameter at discriminating between run types without $WallCut(D)$ (91\% accuracy).}
\end{figure}

@@

Gravitational differentiation is a novel technique for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.

It accomplishes this using a new method for calculating gradients for the last layer of the neural network. Based on the current predictions of a trained network on a certain unlabeled training example, a gradient is calculated for the last layer with the following purposes:

@@

\begin{enumerate}
    \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for $\upsilon$.
    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $GravDiff(p, \psi, g)$ and record the resulting gradients for the next iteration.
    \item Increment $g$ by $\delta_{g}$. At the beginning, when $g = 0$, training will be entirely based on the labeled set $\varsigma$, and progressively, over the course of a run, the gravitational effect increases with $g$.
\end{enumerate}

@@

@@

Iterative cluster nucleation is a semi-supervised learning algorithm that takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.

The basic concept is to start with a relatively small amount of labeled data from 2 classes (nuclear recoils and alpha particles, in this case). They can be visualized as clusters. A neural network is trained on the clusters, and used to run predictions on unlabeled data.

Those examples that, based on the network's predictions, fall firmly into one cluster or the other, are added to that cluster, and are used for training. By repeating this process, the clusters are nucleated (expanded) to gradually include more and more of the data set.

>>>>

However, the lack of decisiveness observed even when accuracy was relatively high implies that, without $WallCut(D)$, this is a suboptimal technique.

@@

Gravitational differentiation is a novel technique, developed during this study, for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.

In general terms, it amplifies the training effect associated with high-confidence predictions (those that most clearly show characteristics of either an alpha particle or a neutron) on unlabeled examples. Meanwhile, it minimizes the training impact of low-confidence predictions, such that the network will not be trained to make incorrect predictions.

Analogous to a gravitational effect, events with high-confidence predictions are attracted closer to their corresponding classifications. They are then used as training data, allowing the patterns they represent to be used more generally to make predictions on other events.

The system accomplishes this using a new method for calculating gradients for the last layer of the neural network. Based on the current predictions of a trained network on a certain unlabeled training example, a gradient is calculated for the last layer with the following purposes:

@@

\begin{enumerate}
    \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for each of the examples in $\upsilon$.
    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $GravDiff(p, \psi, g)$ and record the resulting gradients for the next iteration.
    \item Increment $g$ by $\delta_{g}$. At the beginning, when $g = 0$, training will be entirely based on the labeled set $\varsigma$, and progressively, over the course of a run, the gravitational effect increases with $g$.
\end{enumerate}

@@

!!!Include table on grid search!!!

@@

Iterative cluster nucleation is a second semi-supervised learning algorithm developed as part of this study. It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.

The basic concept is similar to that of gravitational differentiation. Starting with a relatively small amount of labeled data from 2 classes (nuclear recoils and alpha particles, in this case), predictions are run on a set of unlabeled data, and those predictions are used to produce further training data. However, there are 2 key differences:

\begin{enumerate}
    \item In iterative cluster nucleation, an unlabeled example is not included in the training process at all, until a very confident prediction is made on it. Once that occurs, it is added to the training set, expanding 1 of the 2 ``clusters'' of training data.
    \item Once an example is added to the active training set, it is never removed. This ensures that the highly confident predictions made at the very beginning continue to influence the training process.
    \item The unlabeled examples, once added to the active training set, are weighted just as strongly as examples from the original labeled data set. This ensures imperfect labels in the labeled data do not outweigh the confident and more likely correct predictions from later in the training process.
\end{enumerate}
